---
title: Analyzing Top 10.000 Most Popular Movies on IMDB
description: Finding the common characteristics among Top 10.000 movies of all time
execute: 
  echo: false
---

# Executive Summary

## Objective

The goal of this analysis is to find **common characteristics** among the **most popular movies**. These characteristics would be helpful to potential **movie producers** who wants to identify trends and patterns that contribute to a film's success.

## Methodology

### Data Collection

The data was collected from [IMDB Non-Commercial Dataset](https://developer.imdb.com/non-commercial-datasets/), which are freely available to the public.

### Data Cleaning

The dataset is given as a list of `.tsv` files. To process that, I setup a local SQLite database with appropriate tables and imported the `.tsv` into them.

The dataset contains a comprehensive amount of data, with a total of 39 columns that are spread across of 7 tables. For this analysis, I selected 2 tables that are most relevant and joined them together: `title.basics` and `title.ratings.`

To get the most popular movies, I sorted the movies descendingly based on the number of votes, as it should be positively correlated with the amount of people that watched the movie.

Finally, I took the top 10.000 rows and export them into a `.csv` file.

The `csv` file is then loaded into R to check for data quality and for further analysis. The data is already clean. The data reflects what the live site shows, with 0,01% null values, no duplicates, no misspellings, and so on.

### Data Analysis

The analysis is also performed using R in RStudio, with the help of `tidyverse` libraries such as `dplyr` and `ggplot2.` The majority of the analysis is identifying relationships among the number of votes and the potential key variables using statistics and visualization.

## Results

### Most Common Movie Duration

```{r}
df |> 
  ggplot(aes(runtimeMinutes)) +
  geom_histogram(aes(fill = runtimeMinutes >= 80 & runtimeMinutes <= 120), 
                 binwidth = 40, boundary = 0) +
  scale_x_continuous(breaks = seq(0, 240, 40)) +
  scale_fill_manual(values = c("FALSE" = "grey", "TRUE" = "#FFA07A"),
                    labels = c("FALSE" = "Other Durations", "TRUE" = "Most Common Durations")) +
  labs(fill = "",
       title =  "Most Common Movie Duration",
       subtitle = '73% of Movies are between 80-120 minutes long') +
  theme(legend.position = "bottom")
```

### Most Common Genre Combinations

```{r}
df |> 
  count(genres) |> 
  arrange(desc(n)) |>
  mutate(is_top3 = ifelse(row_number() <= 3, "Top 3", "Others")) |>  # Flag top 5
  slice_head(n = 10) |> 
  ggplot(aes(n, genres, fill = is_top3)) +  # Use fill aesthetic to differentiate
  geom_col() +
  scale_fill_manual(values = c("Top 3" = "#FFA07A", "Others" = "grey")) +  # Highlight top 5
  theme_minimal() +
  labs(fill = "Group",
       title = 'Most Common Genre Combinations',
       subtitle = 'Each movie can has multiple genres.\nThe most common combinations are Comedy + Drama + Romance,\nfollowed by Drama, and Comedy + Drama')  # Add legend
```

### Most Common Genres

```{r}
df_unnested_genre |> 
  count(genres) |> 
  arrange(desc(n)) |> 
  slice_head(n = 10) |> 
  mutate(rank = if_else(row_number() <= 3, 'Top_3', 'Others')) |> 
  ggplot(aes(genres, n, fill = rank)) +
  geom_col() +
  scale_fill_manual(values = c('Top_3' = '#FFA07A', 'Others' = 'grey')) +
  coord_flip() +
  labs(title = 'Most Common Individual Genres',
       subtitle = 'If we count each genre individually. This is what we get.\nThe top three genres are Drama, Comedy, and Action')
```

## Recommendations

1.  Aim for a runtime between 80-120 minutes: The data shows that 73% of the most popular movies fall within this range. This suggests audiences prefer films of moderate length.

2.  Focus on key genre combinations: Consider producing films that combine Comedy, Drama, and Romance, as this was the most common genre combination. Alternatively, pure Drama films or Comedy-Drama combinations were also highly popular.

3.  Prioritize Drama, Comedy, and Action genres: When considering individual genres, these three were the most common among popular films. Producers might want to ensure their projects incorporate at least one of these genres.
