---
title: "Case Study: Deep Rock Galactic Missions Data"
description: "Analyzing Deep Rock Galactic Daily Missions Data Over July 2024"
author: "invictus"
date: "2024-07-26T16:05:00"
draft: true
toc: true
toc-depth: 6
---

> Rocks and stones may break my bones, but beer will never hurt me. ***gee***, ***I'm feeling poetic today***!

## introduction

Anyone who plays Deep Rock Galactic (DRG) must've heard of that phrase once. It's my personal favorite.

If you're not familiar with the game, DRG is an FPS Sandbox Generated type of game, where you'll play as a miner that lives on space. You'll be diving 1000km to the underground to accomplish a given mission. These missions are randomly generated (or is it?) every 30 minutes.

\<mission screenshot here\>

Interestingly, these missions are the same for all players around the world, as long as you play on the same season.

For every mission, there's a small chance that it'll come with a mutator that changes that nature of the gameplay. One that most sought after is the 'Double XP' mutator, as you can guess, will double the amount of XP you'll obtain.

XP is very important during the early games, because your level will determine what weapons you can unlock and what upgrades you can purchase.

Therefore, one of the reasons that inspired me to conduct this study is to discover the pattern of 'Double XP' mutator. I want to know whether it's completely random or whether I can find out at which hours it usually appear.

I'm a beginner myself in the game, only have about 40 hours gameplay. So, the result of this study will benefit me very much.

### Acknowledgement

This work could've not been done without the massive effort by [rolfosian](https://github.com/rolfosian/), who has written a script to extract the missions data stored on the game. Additionally, he has a site that displays the current active missions updated in real-time, <https://doublexp.net/>. Feel free to check the source code his [github repo](https://github.com/rolfosian/drgmissions).

### Objective

I haven't run it myself on my PC. Fortunately, I discovered that he stores many of the collected missions in json format on the site. Not long after, I wrote simple [python scripts](https://github.com/invictus1032/drg-mission-parser) to download, parse, and convert them to csv.

It was a lot of fun. Now, I want to rewrite the script in R and then finally analyze the data. The objective is to get a full daily missions data over July 2024. And then hopefully to find pattern on 'Double XP'. Additionally, I'll peek on other insights as well, because, why not?

## Collecting Data

Let's fire up our swiss-army knife, `tidyverse`.

```{r}
library(tidyverse)
```

From the rolfosian's project source code, I discovered that the json files are stored on `https://doublexp.net/static/json/bulkmissions/{yyyy-mm-dd}.json.` So let's set that up as the `base_url`

```{r}
json_base_url <- 'https://doublexp.net/static/json/bulkmissions/'
```

The base URL will be used to generate all the download links for each json file. Since the file names are formatted by ISO dates, it's easy to generate using `lubridate`

```{r}
library(lubridate)
start_date <- ymd('2024-07-01')
end_date <- ymd('2024-07-31')
date_list <- seq(start_date, end_date, by = 'day')
```

As simple as that. Now we have a list of date from July 1 to July 31

```{r}
date_list
```

To convert them into complete download links, we can use `paste0()` to combine them with the base URL.

```{r}
json_urls <- paste0(json_base_url, date_list, '.json')
json_urls |> head(5)
```
