## Setup

```{r}
library(tidyverse)
library(janitor)
library(naniar)
library(scales)
library(summarytools)
library(DataExplorer)
library(explore)
library(corrr)
```

### Options

```{r}
options(scipen = 999)
```

### Load Data

```{r}
df_student_gen_23 <- read_csv('student_list_generation_23.csv')
df_student_gen_24 <- read_csv('student_list_generation_24.csv')
```

## Inspect Data

```{r}
df_student_gen_23
```

```{r}
df_student_gen_23 |> 
  glimpse()
```

```{r}
df_student_gen_24
```

```{r}
df_student_gen_24 |> 
  glimpse()
```

### Understand Data

```{r}

```

### Check Data Quality

#### Duplicate

```{r}
df_student_gen_23 |> 
  get_dupes(Nama, NPM)
```

```{r}
df_student_gen_24 |> 
  get_dupes(Nama, NPM)
```

#### Missing Data

```{r}
df_student_gen_23 |> 
  vis_miss()
```

```{r}
df_student_gen_24 |> 
  vis_miss()
```

#### Inconsistency

```{r}
df_student_gen_23
```

###### Class

```{r}
df_student_gen_23 |> 
  mutate(class_length = str_length(`Kelas Baru`)) |> 
  count(class_length)
```

```{r}
df_student_gen_24 |> 
    mutate(class_length = str_length(Kelas)) |> 
    count(class_length)
```

```{r}
df_student_gen_24 |> 
  distinct(Keterangan)
```

```{r}
df_student_gen_24 |> 
  mutate(Keterangan_split = str_split(Keterangan, '-| ')) |> 
  mutate(Keterangan_length = map_int(Keterangan_split, length)) |> 
  count(Keterangan_length)
```

#### Inaccuracy

### Data Quality Summary

1.  4 Duplicate Rows
2.  No Missing Values
3.  Most classes follow 5 letter pattern, in addition to several special classes that vary
4.  Most class description follow a consistent pattern, in addition to several special classes

Further Action: dedicated page for cleaning the data

## Clean Data

The data is cleaned on ![Clean R](clean_R.html)

```{r}
df_combined_cleaned <- read_csv('combined_cleaned_dataset.csv')
```

## Analyze Data

### Correlation
