---
title: "Case Study: RevoU DAMC Certification (July 22)"
description: "See how I tackle RevoU DAMC's Certification Case Study using R"
date: "2024-07-24T16:05:00"
draft: false
categories: [Fictional Project, R]
---

## Introduction

Recently, I enrolled in a Mini-Course Data Analytics course from RevoU. It's a basic overview and introduction to the world of Data Analytics. At the end of it, you'll get a case study you can work on. That case study is the one I'll be working on today.

Actually, I've done a pretty rough analysis and visualization with it, but it was all over the place. So, I want to redo it from scratch, hopefully in a more structured format.

### About the case study

### Description

The case study contains a dataset about sales data of a fictional company. The columns are described as follows:

> **ORDERNUMBER**: A unique identifier for each sales order.
>
> **QUANTITYORDERED**: The number of units ordered for a particular product.
>
> **PRICEEACH**: The price of a single unit of the product.
>
> **ORDERDATE:** The date when the order was placed.
>
> **STATUS**: The current status of the order (e.g., Shipped, Cancelled, On Hold).
>
> **PRODUCTLINE**: The category or line to which the product belongs.
>
> **PRODUCTCODE**: A unique code identifying the product.
>
> **CUSTOMERNAME**: The name of the customer who placed the order.
>
> **CITY**: The city where the customer is located.
>
> **DEALSIZE**: The size category of the deal (e.g., Small, Medium, Large)

### Task

The task is to find answers to the following questions:

> 1.  Which product lines have the highest and lowest sales? Create a chart that is representable.
> 2.  Show sales performance over time, is there any pattern?
> 3.  How does deal size (small, medium, large) correlate with total sales? What is the percentage of contribution for each type of deal?

## Get Started

### Inspect and Clean Data

Without further ado, let's load the `tidyverse`!

```{r}
library(tidyverse)
```

Just to be safe, let's take look at the first few lines of the csv.

```{r}
read_lines('./revou_damc_case_study.csv', n_max = 5)
```

Looks good enough. Now let's load the file and take a closer look

```{r}
df <- read_csv('./revou_damc_case_study.csv')
```

```{r}
df |> str()
```

```{r}
df
```

#### Clean Date

The first glaring issue is the date. It's still in \<chr\>. However, the first few rows have all the same hms (hours, minutes, seconds). So, there's a possibility it's the same all the way down. Let's see...

```{r}
df |> 
  separate_wider_delim(ORDERDATE, delim = ' ', names = c('ORDERDATE', 'ORDERHOUR')) |> 
  count(ORDERHOUR)
```

I knew it. It's very unlikely for all orders to be exactly on the same hours. So, the most probable thing is they don't have data on the hms, so they just insert 0:00 as a placeholder. Since the placeholder is not useful for us, let's drop it.

```{r}
df <- df |> 
  separate_wider_delim(ORDERDATE, delim = ' ', names = c('ORDERDATE', 'ORDERHOUR')) |> 
  select(!ORDERHOUR)
```

Let's also parse the date while we're at it.

```{r}
df <- df |> 
  mutate(ORDERDATE = mdy(ORDERDATE))
```

`ORDERDATE` is likely be one of the key variables in our analysis. So, we'll frequently work with that

#### Clean Price

Since we're dealing with sales data, the next important thing is the numbers: price, quantity, revenue. Let's take a look again at the data.

```{r}
df
```

Unfortunately, the case study doesn't provide any information on the currency. So, let's assume it's in USD. Now let's see the distribution.

```{r}
df |> 
  ggplot(aes(PRICEEACH)) +
  geom_histogram()
```

The price is indeed unusual. From the first few rows, it even seems that it's not just about ranging around 100, but it's exactly 100. Let's set the `binwidth` to `1` to prove it.

```{r}
df |> 
  ggplot(aes(PRICEEACH)) +
  geom_histogram(binwidth = 1)
```

Yup, exactly. Let's see it from the table for clearer numbers using `count()`

```{r}
df |> 
  count(PRICEEACH, sort = TRUE)
```

Yeah... I'll be honest, I have no idea why this could've happened, unless the data is fake. (Oh wait, it is!). But in a real world, we can't do that assumption. We have to be prepared for any unexpected data. So, let's treat this dataset as a real one even though we know it's not.

Okay, so the data is sales data, precisely orders data. So, what we got previously is the distribution of the sales, not the price of the product themselves. So, one explanation of it happened is because the products that are most sought of are the one that has the price of \$100.

Nevertheless, how many products are there that are priced at \$100? 1? 2? if there's many products priced at \$100, then we can assume there's nothing special with the products. They're just the most common.

Let's see...

```{r}
df |> 
  distinct(PRODUCTCODE, PRICEEACH) |> 
  ggplot(aes(PRICEEACH)) +
  geom_histogram(binwidth = 1)
```

Now the data is getting weirder. Why would a store prices most of the products at \$100?

Again, I have no idea. So, if you know something, let me know in the comments. s

#### Clean QuantityOrdered

The next important variable is `QUANTITYORDERED`. Let's see if it also has similar anomaly

```{r}
df |> 
  ggplot(aes(QUANTITYORDERED)) +
  geom_histogram(binwidth = 1)
```

Finally, a normal data! *just kidding* 😂. Anyway, the distribution looks acceptable. Few outliers there and there, most values are around the median, etc. I'm not familiar with business, but I'm quite certain it doesn't have the same anomaly as the `PRICEEACH`.

### Analyze and Visualize

The important variables has been checked. Now it's time to solve the questions.

#### 1. Product Lines Sales

> 1.  Which product lines have the highest and lowest sales? Create a chart that is representable.

Oh, right. Forgot to check the categorical variables. Welp, let's just deal with it as we go. Anyway, I assume 'sales' here mean the quantity of units sold. We can compute that from aggregating the `PRODUCTLINE` with their `QUANITITYORDERED`.

```{r}
df |> count(PRODUCTLINE, wt = QUANTITYORDERED, sort = TRUE)
```

Ohh, this is why I love `count()`. Anyway, the task is to create a chart. So let's fire up `ggplot`.

```{r}
df |> 
  count(PRODUCTLINE, wt = QUANTITYORDERED) |> 
  ggplot(aes(fct_reorder(PRODUCTLINE, n), n)) +
  geom_col() +
  labs(x='Product Lines', y='Total Sales')
```

From this, we can see that the lowest sales are from `Trains`, and the highest is from `Classic Cars`.

#### 2. Sales Performance Pattern

> 2.  Show sales performance over time, is there any pattern?

Again, we need to understand what does `Sales Perfomance` mean? It can be measured from various metrics, depending on the goals, so we need to guess. Since it's just a simple case study, let's go with revenue, which can be obtained from multiplying `QUANTITYORDERED` and `PRICEEACH`

First, let's create a new `revenue` column

```{r}
df <- df |> 
  mutate(revenue = QUANTITYORDERED * PRICEEACH, .after = PRICEEACH)

df
```

##### By Product Lines

Now, we can compute the total revenue of each `PRODUCTLINE` per day using `count()`

```{r}
daily_revenue <- df |> 
  count(ORDERDATE, PRODUCTLINE, wt = revenue, name = 'REVENUE')

daily_revenue
```

```{r}
daily_revenue |> 
  ggplot(aes(ORDERDATE, REVENUE, color=PRODUCTLINE)) +
  geom_line()
```

Wow, there's too much lines on the same place. Let's divide them up with `facet`.

```{r}
daily_revenue |> 
  ggplot(aes(ORDERDATE, REVENUE, color=PRODUCTLINE)) +
  geom_line() +
  facet_wrap(. ~ PRODUCTLINE, ncol = 1)
```

It looks better now, still feel too much. The Y axis is barely readeable. Maybe we need to display fewer `PRODUCTLINE` at a time.

##### All Sales

Actually, I just realized that it wasn't asked to group by `PRODUCTLINE`. So, we could've just plot the entire sales. For this, we need to re-aggregate the `daily_revenue` to get a summarized revenue across all `PRODUCTLINE`.

```{r}
daily_revenue_total <- daily_revenue |> 
  count(ORDERDATE, wt = REVENUE, name = 'total_revenue')
```

```{r}
daily_revenue_total |> 
  ggplot(aes(ORDERDATE, total_revenue)) +
  geom_line() 
```

We can roughly see the peaks in the later parts of the year. We can add a `geom_smooth()` to help visualize the pattern.

```{r}
daily_revenue_total |> 
  ggplot(aes(ORDERDATE, total_revenue)) +
  geom_line() +
  geom_smooth()
```

Maybe we can see the pattern better if we divide the plot by year

```{r}
daily_revenue_total <- daily_revenue_total |> 
  mutate(year = year(ORDERDATE) |> as_factor())
```

```{r}
daily_revenue_total |> 
  ggplot(aes(ORDERDATE, total_revenue)) +
  geom_line() +
  facet_wrap(. ~ year, scales = 'free_x', ncol = 1) +
  scale_x_date(date_labels = '%b',date_breaks = '1 month')
```

Now we got a clearer broad overview of the trend by year. However, the small size of the plot make it a little bit hard to identify the pattern. Let's zoom in by drawing each plot for each year.

```{r}
daily_revenue_total |> 
  filter(year(ORDERDATE) == 2003) |> 
  ggplot(aes(ORDERDATE, total_revenue)) +
  geom_line() + 
  geom_smooth() +
  scale_x_date(date_labels = '%b',date_breaks = '1 month') +
  labs(title = 'Total Sales Revenue per Month in 2003')
```

Now we can clearly see the peak on around November - December. Besides, the sales seem to be already going upward starting from September

```{r}
daily_revenue_total |> 
  filter(year(ORDERDATE) == 2004) |> 
  ggplot(aes(ORDERDATE, total_revenue)) +
  geom_line() + 
  geom_smooth() +
  scale_x_date(date_labels = '%b',date_breaks = '1 month') +
  labs(title = 'Total Sales Revenue per Month in 2004')
```

In 2005, the pattern a bit fluctuates. Yet, there's still a generate upward trend toward the last months of the year. Moreover, the peaks are the same, which are around November - December

```{r}
daily_revenue_total |> 
  filter(year(ORDERDATE) == 2005) |> 
  ggplot(aes(ORDERDATE, total_revenue)) +
  geom_line() + 
  geom_smooth() +
  scale_x_date(date_labels = '%b',date_breaks = '1 month') +
  labs(title = 'Total Sales Revenue per Month in 2005')
```

Unfortunately, in 2005, we're lacking data toward the later months. So we can't see the peaks on the same months.

However, based on the pattern of the previous 2 years, we can predict that sales will also increase upward toward the last part of the years, and peak around November - December
